\sec{Introduction}
This file has been made to serve as a compilation of methods taught in the course MA 108.\\
The main purpose for making this file is the following - given the scenario regarding the abrupt end of this (Spring 2020) semester, this course had to be abandoned. However, this is a useful course for many and thus, one must at least be aware of the basics.\\
While this cannot serve as a complete substitute for the course in terms of theory, it's the next best thing to have. Ignoring the finer details of when certain things exist (converge), this document covers the things that one should remember from this course.\\
Due to this, I also skip proofs and derivations for the most part. For these, I refer the following two sources:
\begin{enumerate}[leftmargin=*, nosep]
	\item Professor Preeti Raman's page for this course - \url{http://www.math.iitb.ac.in/~preeti/ma108-2019/}.
	\item Professor Gopala K. Srinivasan's notes for this course - \url{http://www.math.iitb.ac.in/~gopal/MA108/ma108_handwritten_notes_2008.pdf}.
\end{enumerate}

These are also the references I have used for making this document.\\
Few things are written in a red box. These can be ignored. The last section can also be ignored. \\
Also, I have skipped the extremely basic \cancel{JEE} things about classification of ODEs and solving homogeneous ODEs. These can be found at the beginning of the above notes.

\begin{mdframed}[style=boxstyle2, frametitle={A note on abuse of notation}]
	In this document, the following abuse of notation is often done - I write something like ``the function $f(t)$'' or ``the function $u(x, y)$.''\\
	The correct thing to write is ``the function $f$'' or ``the function $u$.''\\
	However, sometimes it is more illustrating to abuse notation.
\end{mdframed}

These notes have not been thoroughly proofread and thus, it is possible that there are misteaks\footnote{intended} (mostly typos). If you find any, do let me know.\\
You can anonymously send a feedback here - \\
\url{https://forms.gle/nif2qPuB7GfSbqxt5}.