\sec{The Method of Variation of Parameters}
\subsection{Introduction}
This is a powerful method to find a particular solution to a linear ODE when the general solution of the associated homogeneous equation is known.
\begin{mdframed}[style=boxstyle, frametitle={The Setup}]
	Consider the second order ODE 
	\begin{equation} \label{eq:mvp}
		y'' + P(x)y' + Q(x)y = R(x).
	\end{equation}
	Assume that the general solution $y_g(x)$ of the associated homogeneous equation
	\begin{equation} \label{eq:mvphomo}
		y'' + P(x)y' + Q(x)y = 0
	\end{equation}
	is known.
\end{mdframed}
\begin{mdframed}[style=boxstyle2, frametitle={The Solution - derivation. (Can be ignored.)}]
	Let $y_1$ and $y_2$ be two linearly independent solutions of (\ref{eq:mvphomo}).\\
	We seek to find a particular solution $y_p(x)$ of (\ref{eq:mvp}) of the form
	\begin{equation} \label{eq:ypform}
		y_p(x) = v_1(x)y_1(x) + v_2(x)y_2(x).
	\end{equation}
	(It is our \emph{assumption - I} that we would have a solution of this form.)\\
	Now, we place the following restriction
	\begin{equation} \label{eq:mvpcond1}
		v_1'y_1 + v_2'y_2 \equiv 0.
	\end{equation}
	(This is \emph{assumption - II}.)\\
	Under this assumption, we get 
	\begin{align*} 
		y_p' &= v_1y_1' + v_2y_2',\\
		y_p'' &= v_1y_1'' + v_1'y_1' + v_2y_2''+ v_2'y_2'.
	\end{align*}
	Substituting this back into (\ref{eq:mvp}) and using the fact that $y_1$ and $y_2$ are solutions of (\ref{eq:mvphomo}) gives us that
	\begin{equation} \label{eq:mvpcond2}
		v_1'y_1' + v_2'y_2' = R(x).
	\end{equation}
	Now, note that the system of equations (\ref{eq:mvpcond1}) and (\ref{eq:mvpcond2}) can be uniquely solved for $v_1'$ and $v_2'$. (Since $\{y_1, y_2\}$ is linearly independent and hence, the Wronskian is always nonzero. Recall Corollary \ref{cor:wronskian}.)\\
	Integrate to obtain $v_1$ and $v_2$ and finally, a particular solution $y_p$ is given by (\ref{eq:ypform}). The general solution is, as usual, $y_g + y_p.$
\end{mdframed}
\begin{mdframed}[style=boxstyle, frametitle={The Solution}]
	Let $y_1$ and $y_2$ be two linearly independent solutions of (\ref{eq:mvphomo}).\\
	Solve the following system of equations for functions $v_1'(x)$ and $v_2'(x)$:
	\begin{align*} 
		v_1'y_1 + v_2'y_2 &= 0,\\
		v_1'y_1' + v_2'y_2' &= R(x).
	\end{align*}
	Integrate $v_1'$ and $v_2'$ to get $v_1$ and $v_2.$ Then, a particular solution of (\ref{eq:mvp}) is:
	\begin{equation*} 
		y_p(x) = v_1(x)y_1(x) + v_2(x)y_2(x).
	\end{equation*}
	The general solution of (\ref{eq:mvp}) is then $y_g + y_p.$
\end{mdframed}
\exercise{%
Solve $y'' + y = \tan x.$
}