\begin{rubric}{Reading Projects}

\entry*[2021*] \textbf{Homological Algebra} \hfill \emph{Guide: Prof. Dipendra Prasad} \,|\, \emph{IIT Bombay}

	Reading \emph{An Introduction to Homological Algebra} by \emph{Charles A. Weibel} and have covered topics such as Abelian categories, Chain complexes, Chain homotopies, $\delta$-functors, Derived functors, Projective and Injective resolutions, Adjoint functors, Balancing Tor and Ext, Tor for abelian groups, Tor and Flatness, Ext and Extensions, Derived Functors of the Inverse Limit.

\entry*[2021] \textbf{Algebraic Geometry} \hfill \emph{Guide: Prof. Arvind Nair} \,|\, \emph{TIFR}

	Participated in the month long Visiting Students' Research Programme conducted by Tata Institute of Fundamental Research. 

	Read \emph{Foundations of Algebraic Geometry} by \emph{Ravi Vakil} and covered topics such as Sheaves, Affine schemes, Morphisms of Schemes	and Chevalley's Theorem. 

	The final presentation can be found \href{https://aryamanmaithani.github.io/math/alg-geo/}{here}.

\entry*[2021] \textbf{Representation Theory} \hfill \emph{Guide: Prof. Ananthnarayan Hariharan} \,|\, \emph{IIT Bombay}
	
	Read \emph{Representation Theory of Finite Groups} by \emph{Benjamin Steinberg} and covered topics such as Maschke's Theorem, the orthogonality relations, characters, permutations and regular representations, induction of representations, dimension theorem, representations of dihedral and symmetric groups and applications to group theory such as Burnside's $pq$-theorem. 

	The report can be found \href{https://aryamanmaithani.github.io/math/rep-theory/Notes.pdf}{here}.

\entry*[2020] \textbf{Algebraic Topology}  \hfill \emph{Guide: Prof. Rekha Santhanam} \,|\, \emph{IIT Bombay}

	Read homotopy theory including theorems like Van Kampert Theorem, existence of a universal covering space, Galois correspondence for covering spaces and applications to group theory.

	Also read homology theory and applications such as invariance of domain, degree of maps, hairy ball theorem as well as CW complexes.

	The report and related presentation can be found \href{https://aryamanmaithani.github.io/math/alg-top/}{here}.

\entry*[2020] \textbf{Primes is in P} \hfill \emph{Guide: Prof. Ronnie Sebastian} \,|\, \emph{IIT Bombay}

	Read the paper \emph{PRIMES is in P} by \emph{Manindra Agrawal, Neeraj Kayal, Nitin Saxena} about the AKS Algorithm and wrote a report on that, which can be found at the end of notes \href{http://www.math.iitb.ac.in/~ronnie/Spring2020/Lecture-Notes-Galois-Theory.pdf}{here}.


\entry*[2019] \textbf{Posets} \hfill \emph{Guide: Prof. Koushik Saha} \,|\, \emph{IIT Bombay}

	Read about posets from \emph{Enumerative Combinatorics} by \emph{Richard P. Stanley} and gave a presentation on it as part of extra reading in Combinatorics, which can be found \href{https://aryamanmaithani.github.io/math/posets/}{here}.

\entry*[2019] \textbf{Topology - Classification of Surfaces} \hfill \emph{Guide: Senior Student} \,|\, \emph{IIT Bombay}

	Read about Topology and made a report that contains an introduction to topology and a proof of the Classification of Surfaces, which can be found \href{https://aryamanmaithani.github.io/math/classification-of-surfaces/Classification%20of%20Surfaces.pdf}{here}.
\end{rubric}